Extreme heterogeneity of influenza virus transcription in single cells

### 2017-09-12 paper/paper.tex
Tweaks to Methods

\begin{figure}[t!]
\centerline{\includegraphics[width=0.9\linewidth]{figures/p_frac_flu_summary.pdf}}
\caption{\label{fig:viralbarcodes}
Synonymous barcodes on the viral mRNAs distinguish true infections from cells that contain viral mRNAs derived from leakage of lysed cells.
{\bf (A)}
Cells with at least two viral mRNAs for which the barcode could be called, arranged in order of increasing influenza transcript counts.
Bar heights denote the number viral mRNAs on a log\textsubscript{10} scale, bar coloring is linearly proportional to the fractions of viral mRNAs derived from wild-type and synonymously barcoded virus.
{\bf (B)}
Same as (A), but each bar is colored according to the relative fraction of the more common (major) and less common (minor) virus variant.
At low levels of viral mRNA there is often a roughly equal mix, suggesting contamination with viral mRNAs leaked from lysed cells.
At higher levels of viral mRNA, cells generally have only one viral variant, suggesting infection initiated by a single particle.
A few cells are also obviously co-infected with both viral variants.
{\bf (C)}
We determined a cutoff for calling ``true'' infections by fitting a curve to the mean barcode purity of all cells with greater than a given fraction of their mRNA derived from virus.
We called the cutoff at the point at which purity stops increasing with the fraction of viral mRNA.
{\bf (D)}
The number of cells identified as infected and co-infected for each sample, as well as the number of cells with any viral read.
For all subsequent analyses, we subsampled the number of uninfected cells per sample to the greater of 50 or the number of infected cells.
{\bf (E)} 
Distribution of the fraction of mRNA per cell derived from virus for both infected and co-infected cells.
}
\figsupp[Number of viral barcodes called.\label{figsupp:barcodescalled}]{
The number of viral barcodes called for each sample and influenza gene segment. 
Viral transcripts are classified as \emph{syn} if they mapped to a synonymously barcoded influenza transcript, \emph{wt} if they mapped to a wild-type influenza transcript, \emph{invalid} if multiple reads for the same UMI differed on the status of the viral barcode, and as \emph{uncalled} if none of the reads for that UMI overlapped the region of the viral transcript containing the viral barcode.
}{\includegraphics[width=\linewidth]{figures/synbarcodes_umistats.jpg}}
\figsupp[Fraction of total viral mRNA derived from a given fraction of infected cells.\label{figsupp:cumulfracflu}]{
The total fraction of all viral mRNA among infected cells that is attributable to a given fraction of these cells.
For instance, the plot for the 8-hour sample shows that roughly 50\% of all viral mRNA is derived from 10\% of the infected cells.
}{\includegraphics[width=\linewidth]{figures/p_cumul_flu.pdf}}
\end{figure}

@@

\begin{figure}[t!]
\centerline{\includegraphics[width=0.9\linewidth]{figures/p_flu_burden_flu_gene_merge.pdf}}
\caption{\label{fig:fluburdenbyflugene}
The absence of viral genes explains some of the variability in the amount of viral mRNA per cell.
{\bf (A)} 
Fraction of mRNA in each infected cell derived from virus as a function of the normalized expression of each viral gene in that cell, taken over all time points.
Cells with high viral burden always express all RNP genes, but some cells with high viral burden lack each of the other genes.
{\bf (B)}
Box plots showing the per-cell viral burden among cells with $>$0.5\% of their mRNA from virus, binned by whether or not the cells express each gene.
A Wilcoxon signed-rank test was used to test the null hypothesis that absence of each gene does not affect viral burden: **** = $P < 10^{-4}$, *** = $P < 10^{-3}$,  * = $P < 0.05$, ns = not significant.
Similar results are obtained if we examine only the 10-hour timepoint (Figure~\ref{fig:fluburdenbyflugene}-Figure~supplement~\ref{figsupp:10hrfluburdenbyflugene}).
Top and bottom of box bound the first and third quartiles; the line in the box is at the median.
Whiskers extend to the highest or lowest point within1.5$\times$ the interquartile range, and outliers are shown as circles.
{\bf (C)}
The fraction of that express each of the four other genes among cells that express all RNP genes, as well as the fraction that express \emph{all} four of the other genes.
The fraction that express all four genes is well predicted by simply multiplying the frequencies of cells that express each gene individually, indicating that gene absence is approximately independent across these genes.
}
\figsupp[Secondary transcription from newly synthesized RNPs is a major source of viral mRNA during bulk infections.\label{figsupp:cyclohexamide}]{
A549 cells were infected at an MOI of 0.2 as calculated on MDCK-SIAT1 cels in either the presence or absence of the protein-translation inhibitor cyclohexamide, and viral mRNA was quantified at 8 hours post-infection by qPCR.
The cyclohexamide prevents translation of new PB2, PB1, PA, and NP protein, and so prevents the formation of the new RNPs needed for secondary transcription.
The bars show the relative amount of HA and PB2 mRNA in the absence versus the presence of cyclohexamide. Error $\pm$ S.D. n=3. 
}{\includegraphics[width=0.3\linewidth]{figures/Primary_transcription_measurements/Cyclohexamide.pdf}}
\figsupp[Like panel (B) but for the 10-hr sample only.\label{figsupp:10hrfluburdenbyflugene}]{
The absence of viral RNP genes but \emph{not} non-RNP genes remains significantly associated with reduced viral burden when we examine only the 10-hr sample, which is the single time point with the most data points.
The difference for NP is no longer statistically significant due to low counts of infected cells lacking NP, but the trend remains.
}{\includegraphics[width=\linewidth]{figures/p_10hr_flu_burden_flu_gene_test}}
\figdata{The numerical data for panel (C) are in \texttt{p\_missing\_genes.csv}.}
\end{figure}

@@

Wildtype A/WSN/1933 (H1N1) influenza virus was generated by reverse genetics using the plasmids pHW181-PB2, pHW182-PB1, pHW183-PA, pHW184-HA, pHW185-NP, pHW186-NA, pHW187-M, and pHW188-NS~\citep{hoffmann2000dna}.
The sequences of the viral RNAs encoded in these plasmids are in Figure~\ref{fig:workflow}-source~data~\ref{figdata:viralsequences}.
Reverse-genetics plasmids encoding the synonymously barcoded WSN virus were created by using site-directed mutagenesis to introduce two synonymous mutations near the 3' end of the mRNA for each viral gene.
The sequences of the synonymously barcoded viral RNAs are in Figure~\ref{fig:workflow}-source~data~\ref{figdata:viralsequences}
To generate viruses from these plasmids, we transfected an equimolar mix of all eight plasmids into cocultures of 293T and MDCK-SIAT1 cells seeded at a ratio of 8:1. 
At 24 hours post-transfection, we changed media from D10 to influenza growth media (Opti-MEM supplemented with 0.01\% heat-inactivated FBS, 0.3\% BSA, 100 U of penicillin/ml, 100  \si{\micro}g of streptomycin/ml, and 100 \si{\micro}g of calcium chloride/ml).
At 48 hours post-transfection we harvested the virus-containing supernatant, pelletted cellular material by centrifugation at 300 x g's for 4 minutes, and stored aliquots at of the clarified viral supernatant -80 \si{\degreeCelsius }. 
We then titered thawed aliquots of viral by TCID50 on MDCK-SIAT1 cells, computing titers via the formula of \citet{reed1938simple}.
To generate our ``high-purity'' stocks of viruses for the single-cell sequencing experiments, we then infected MDCK-SIAT1 cells at an MOI of 0.01, and let the virus replicate for 36 hours prior to harvesting aliquots that were again clarified by low-speed centrifugation, aliquoted, stored at -80  \si{\degreeCelsius }, and titered by TCID50.
The high-MOI passage (high-defective particle) stock used in Figure~\ref{fig:viruspopulations} was generated by instead passaging in MDCK-SIAT1 cells twice at an MOI of 1 for 48 hours, when obvious cytopathic effect was noted.

For the experiments in Figure~\ref{fig:coexpression}-Figure~supplement~\ref{figsupp:flowcyto}, we created a virus that carried an HA gene segment in which the carried GFP in place of most of the HA coding sequence by following a scheme first described by \citet{marsh2007specific}.
Briefly, we created a plasmid encoding a viral RNA with GFP in place of the HA coding sequence in the context of the pHH21~\citep{Neumann:1999ws} reverse-genetics plasmid, removing potential start codons upstream of the GFP (see Figure~\ref{fig:coexpression}-source~data~\ref{figdata:HAflank} for the sequence of the viral RNA)
We then generated GFP-carrying virus by reverse-genetics in previously described cells constitutively expressing HA~\citep{Doud:2016gm}.
A unidirectional vector for generating viral vRNA encoding eGFP flanked by HA packaging sequences was generated from a previously described vector, pHH21~\citep{Neumann:1999ws}. 
To obtain sufficient titers, this HA-eGFP virus was expanded for 44 than than 36 hours after initiating infection at an MOI of 0.01.

\subsection{qPCR}
For the qPCR in Figure~\ref{fig:viruspopulations} and Figure~\ref{fig:fluburdenbyflugene}-Figure~supplement~\ref{figsupp:cyclohexamide}, A549 cells were seeded at 3$\times$10\textsuperscript{5} cells / well in a 6-well tissue culture plate in D10 the day prior to infection. 
On the day of infection, a single well was trypsinized and the cells were counted in order to determine the appropriate amount of virus to use to achieve the intended MOI.
Immediately before infection, D10 was replaced with influenza growth media.
For cells incubated with cyclohexamide, the compound was added to a final concentration of 50  \si{\micro}g/ml at the time of infection -- previously confirmed to be sufficient to halt viral protein production~\citep{Killip:2014kz}.
RNA was purified using the QIAGEN RNeasy plus mini kit following manufacturer's instructions. 
cDNA was synthesized using an oligoDT primer and the SuperScript\textsuperscript{TM} III first-strand synthesis supermix from ThermoFisher using manufacturer's directions. 
Transcript abundance was measured using SYBR\textsuperscript{TM} green PCR master mix, using a combined anneal/extension step of 60 \si{\degreeCelsius } for one minute with the following primers: \emph{HA}: 5'-GGCCCAACCACACATTCAAC-3', 5'-GCTCATCACTGCTAGACGGG-3', \emph{IFNB1}: 5'-AAACTCATGAGCAGTCTGCA-3', 5'-AGGAGATCTTCAGTTTCGGAGG-3', \emph{L32}: 5'-AGCTCCCAAAAATAGACGCAC-3', 5'-TTCATAGCAGTAGGCACAAAGG-3'.   
Biological triplicates were performed for all samples.
For measurements of viral genomic HA content, vRNA was harvested from 80 \si{\micro}l of viral supernatant by the addition of 600 \si{\micro}l of RLT plus before proceeding with the standard QIAGEN Rneasy plus miniprep kit protocol, and cDNA was generated using  SuperScript\textsuperscript{TM} III first-strand synthesis supermix with primers specific for vRNA adapted from \citep{Hoffmann:2001vj}  using the manufacturer's protocol for random-hexamer priming~\citep{Xue:2017dl}.
qPCR was then performed as for mRNA measurements.
In order to obtain molecular resolution on our measurements of HA vRNA, a standard curve was generated from three independent dilutions of the HA-encoding reverse genetics plasmid. 
All vRNA values represent three independent RNA extractions with two replicate qPCR measurements.

\subsection{Flow cytometry titering and analyses}
To determine viral titers in terms of HA-expressing units and for the flow cytometry in Figure~\ref{fig:coexpression}-Figure~supplement~\ref{figsupp:flowcyto}, A549 cells were seeded in a 6-well plate and infected as described above for the qPCR analyses.
Cells were harvested by trypsinization, resuspended in phosphate-buffered saline supplemented with 2\% heat-inactivated FBS, and stained with 10 \si{\micro}g/ml of H17-L19, a mouse monoclonal antibody confirmed to bind to WSN HA in a prior study~\citep{Doud:2017bw}.
After subsequent washes in supplemented PBS, cells were stained with a goat anti-mouse IgG antibody conjugated to APC.
Cells were then washed, fixed in 1\% formaldehyde, and washed further before a final resuspension and analysis. 
To determine HA-expressing units, we then determined the fraction of cells that were HA positive and calculated the HA-expressing units from these fractions assuming a Poisson infection process.

For Figure~\ref{fig:coexpression}-Figure~supplement~\ref{figsupp:flowcyto}, after gating to exclude multiplets in FlowJo, data were subsequently extracted using the R package flowCore~\citep{LeMeur:2007uo} and analyzed using a custom Python script.
Guassian kernel density estimates were obtained using the scipy stats package method, guassian\_kde, using automatic bandwidth determination~\citep{vanderWalt:2017dp}.

\subsection{Infections for single-cell mRNA sequencing}
Single cell sequencing libraries were generated using the 10x Chromium Single Cell 3' platform, V1~\cite{zheng2017massively}.
All time points except for the 8-hour repeat, were performed on the same day.

For the infections, A549 cells were seeded in a 6-well plate, with two wells per time point. 
A single well of cells was trypsinized and counted prior to initiation of the experiment for the purposes of calculating MOI.
Wild-type and synonymously barcoded virus were mixed to an estimated ratio of 1:1 based on prior, exploratory, single-cell analyses (data not shown). 
At the initiation of our experiment, the wells for all time points were changed from D10 to influenza growth media.
Cells were then infected with 0.3 HA-expressing units of virus per cell (a determined by flow cytometry).
The infections were performed in order of time point: first the 10-hour time point, then the 8-hour, and then the 6-hour time point.
At one hour after infection, the media for each time point was changed to fresh influenza growth media.
Note that the HA-expressing units were calculated without this additional washing step, and so likely represent an overestimate of our final infectious dose (consistent with the fact that fewer than 30\% of cells appear infected in the single-cell sequencing data).
All cells were then harvested for single-cell analysis concurrently -- ensuring all had spent equivalent time in changed media .
For our replicate 8-hour time point, cells were infected as above except that the cells were infected at 0.1 HA-expressing units of virus per cell but no wash step was performed, and the sample was prepared on a different day.
After harvest, cells were counted using disposable hemocytometers and diluted to equivalent concentrations with an intended capture of 3000 cells/sample following the manufacturer's provided by 10x Genomics for the Chromium Single Cell platform.
All subsequent steps through library preparation followed the manufacturer's protocol.
Samples were sequenced on an Illumina HiSeq. 
The deep sequencing data are available on the Sequence Read Archive under accession \jdbcomment{Jesse will upload and add this.}

\subsection{Computational analysis of single-cell mRNA sequencing data}
Jupyter notebooks that perform all of the computational analyses are available at \url{https://github.com/jbloomlab/flu_single_cell} \jdbcomment{This GitHub repository will be made public upon acceptance of the manuscript. If you are a reviewer who needs access, please contact the editor.}

Briefly, the raw deep sequencing data were processed using the 10X Genomics software package CellRanger (version 2.0.0).
The reads were aligned to a concatenation of the human and influenza virus transcriptomes.
The human transcriptome was generated by filtering genome assembly GRCh38 for protein coding genes defined in the GTF file GRCh38.87.
The influenza virus transcriptome was generated from the reverse-complement of the wildtype WSN viral RNA sequences as encoded in the reverse-genetics plasmids (Figure~\ref{fig:workflow}-source~data~\ref{figdata:viralsequences}).
CellRanger calls cells based on the number of observed cell barcodes, and creates a cell-gene matrix.
We used custom Python code to annotate the cells in this matrix by the number of viral reads that could be assigned to the wildtype and synonymously barcoded virus.
This annotated cell-gene matrix is in Supplementary~file~\ref{suppfile:cellgenematrix}.
The Jupyter notebook that performs these analyses is at \url{https://github.com/jbloomlab/flu_single_cell/blob/master/align_and_annotate.ipynb} \jdbcomment{This GitHub repository will be made public upon acceptance of the manuscript. If you are a reviewer who needs access, please contact the editor.}, and is also available as Supplementary~file~\ref{suppfile:alignandannotate}.

The annotated cell-gene matrix was then further analyzed in R, primarily using the Monocle package (version 2.4.0)~\citep{Qiu:2017dw,Trapnell:2014kg}.
The Jupyter notebook that performs these analyses is at \url{https://github.com/jbloomlab/flu_single_cell/blob/master/monocle_analysis.ipynb} \jdbcomment{This GitHub repository will be made public upon acceptance of the manuscript. If you are a reviewer who needs access, please contact the editor.}, and is also available as Supplementary~file~\ref{suppfile:monocle}.

Briefly, this analysis was performed as follows.
For each sample, cell barcodes with 2.5-fold fewer or more UMI counts mapping to cellular transcripts than the sample mean were considered to be potential doublets or inefficient/empty reactions, and were excluded from downstream analyses (see red vertical lines in Figure~\ref{fig:cells}C).
In order to determine an appropriate cutoff for how many reads a cell needed to contain in order to be classified as infected, we calculated the mean viral barcode purity across all cells that contained at least a given fraction of viral mRNA and had multiple viral reads that could be assigned a barcode (Figure~\ref{fig:viralbarcodes}B,C).
We then determined the threshold fraction of viral mRNA at which the mean purity no longer increased as a function of the amount of viral mRNA.
This threshold represents the point at which we have effectively eliminated cells that have low barcode purity simply due to lysis-acquired reads sampled randomly from both viral barcodes.
The 10-hour sample has the most viral mRNA, and is expected to have the most lysis-associated influenza reads.
We therefore determined the threshold using just this sample (Figure~\ref{fig:viralbarcodes}C), and then applied this threshold to all samples.
This procedure is expected to be conservative, and may miss some truly infected cells with very low amounts of viral mRNA.
For subsequent analyses, we retained all infected cells and a subsample of uninfected cells (the greater of 50 or the number of infected cells for that sample).
The rationale for subsampling the uninfected cell is that the vast majority of cells are uninfected, and we did not want these cells to completely dominate the t-SNE clustering and other downstream analyses.

>>>>

\begin{figure}[t!]
\centerline{\includegraphics[width=0.9\linewidth]{figures/p_frac_flu_summary.pdf}}
\caption{\label{fig:viralbarcodes}
Synonymous barcodes on the viral mRNAs distinguish true infections from cells that contain viral mRNAs derived from leakage of lysed cells.
{\bf (A)}
Cells with at least two viral mRNAs for which the barcode could be called, arranged in order of increasing influenza transcript counts.
Bar heights denote the number viral mRNAs on a log\textsubscript{10} scale, bar coloring is linearly proportional to the fractions of viral mRNAs derived from wild-type and synonymously barcoded virus.
{\bf (B)}
Same as (A), but each bar is colored according to the relative fraction of the more common (major) and less common (minor) virus variant.
At low levels of viral mRNA there is often a roughly equal mix, suggesting contamination with viral mRNAs leaked from lysed cells.
At higher levels of viral mRNA, cells generally have only one viral variant, suggesting infection initiated by a single particle.
A few cells are also obviously co-infected with both viral variants.
{\bf (C)}
We determined a cutoff for calling ``true'' infections by fitting a curve to the mean barcode purity of all cells with greater than a given fraction of their mRNA derived from virus.
We called the cutoff at the point at which purity stops increasing with the fraction of viral mRNA.
{\bf (D)}
The number of cells identified as infected and co-infected for each sample, as well as the number of cells with any viral read.
For all subsequent analyses, we subsampled the number of uninfected cells per sample to the greater of 50 or the number of infected cells.
{\bf (E)} 
Distribution of the fraction of mRNA per cell derived from virus for both infected and co-infected cells.
}
\figsupp[Number of viral barcodes called.\label{figsupp:barcodescalled}]{
The number of viral barcodes called for each sample and gene segment. 
Viral transcripts are classified as \emph{syn} if they mapped to a synonymously barcoded influenza transcript, \emph{wt} if they mapped to a wild-type influenza transcript, \emph{invalid} if multiple reads for the same UMI differed on the status of the viral barcode, and as \emph{uncalled} if none of the reads for that UMI overlapped the region of the viral transcript containing the viral barcode.
}{\includegraphics[width=\linewidth]{figures/synbarcodes_umistats.jpg}}
\figsupp[Fraction of total viral mRNA derived from a given fraction of infected cells.\label{figsupp:cumulfracflu}]{
The total fraction of all viral mRNA among infected cells that is attributable to a given fraction of these cells.
For instance, the plot for the 8-hour sample shows that roughly 50\% of all viral mRNA is derived from 10\% of the infected cells.
}{\includegraphics[width=\linewidth]{figures/p_cumul_flu.pdf}}
\end{figure}

@@

\begin{figure}[t!]
\centerline{\includegraphics[width=0.9\linewidth]{figures/p_flu_burden_flu_gene_merge.pdf}}
\caption{\label{fig:fluburdenbyflugene}
The absence of viral genes explains some of the variability in the amount of viral mRNA per cell.
{\bf (A)} 
Fraction of mRNA in each infected cell derived from virus as a function of the normalized expression of each viral gene in that cell, taken over all time points.
Cells with high viral burden always express all RNP genes, but some cells with high viral burden lack each of the other genes.
{\bf (B)}
Box plots showing the per-cell viral burden among cells with $>$0.5\% of their mRNA from virus, binned by whether or not the cells express each gene.
A Wilcoxon signed-rank test was used to test the null hypothesis that absence of each gene does not affect viral burden: **** = $P < 10^{-4}$, *** = $P < 10^{-3}$,  * = $P < 0.05$, ns = not significant.
Similar results are obtained if we examine only the 10-hour timepoint (Figure~\ref{fig:fluburdenbyflugene}-Figure~supplement~\ref{figsupp:10hrfluburdenbyflugene}).
Top and bottom of box bound the first and third quartiles; the line in the box is at the median.
Whiskers extend to the highest or lowest point within1.5$\times$ the interquartile range, and outliers are shown as circles.
{\bf (C)}
The fraction of that express each of the four other genes among cells that express all RNP genes, as well as the fraction that express \emph{all} four of the other genes.
The fraction that express all four genes is well predicted by simply multiplying the frequencies of cells that express each gene individually, indicating that gene absence is approximately independent across these genes.
}
\figsupp[Secondary transcription from newly synthesized RNPs is a major source of viral mRNA during bulk infections.\label{figsupp:cyclohexamide}]{
A549 cells were infected at an MOI of 0.2 as calculated on MDCK-SIAT1 cells in either the presence or absence of the protein-translation inhibitor cyclohexamide, and viral mRNA was quantified at 8 hours post-infection by qPCR.
The cyclohexamide prevents translation of new PB2, PB1, PA, and NP protein, and so prevents the formation of the new RNPs needed for secondary transcription.
The bars show the relative amount of HA and PB2 mRNA in the absence versus the presence of cyclohexamide. Error $\pm$ S.D. n=3. 
}{\includegraphics[width=0.3\linewidth]{figures/Primary_transcription_measurements/Cyclohexamide.pdf}}
\figsupp[Like panel (B) but for the 10-hr sample only.\label{figsupp:10hrfluburdenbyflugene}]{
The absence of viral RNP genes but \emph{not} non-RNP genes remains significantly associated with reduced viral burden when we examine only the 10-hr sample, which is the single time point with the most data points.
The difference for NP is no longer statistically significant due to low counts of infected cells lacking NP, but the trend remains.
}{\includegraphics[width=\linewidth]{figures/p_10hr_flu_burden_flu_gene_test}}
\figdata{The numerical data for panel (C) are in \texttt{p\_missing\_genes.csv}.}
\end{figure}

@@

Wildtype A/WSN/1933 (H1N1) influenza virus was generated by reverse genetics using the plasmids pHW181-PB2, pHW182-PB1, pHW183-PA, pHW184-HA, pHW185-NP, pHW186-NA, pHW187-M, and pHW188-NS~\citep{hoffmann2000dna}.
The sequences of the viral RNAs encoded in these plasmids are in Figure~\ref{fig:workflow}-source~data~\ref{figdata:viralsequences}.
Reverse-genetics plasmids encoding the synonymously barcoded WSN virus were created by using site-directed mutagenesis to introduce two synonymous mutations near the 3' end of the mRNA for each viral gene.
The sequences of the synonymously barcoded viral RNAs are in Figure~\ref{fig:workflow}-source~data~\ref{figdata:viralsequences}.

To generate viruses from these plasmids, we transfected an equimolar mix of all eight plasmids into cocultures of 293T and MDCK-SIAT1 cells seeded at a ratio of 8:1. 
At 24 hours post-transfection, we changed media from D10 to influenza growth media (Opti-MEM supplemented with 0.01\% heat-inactivated FBS, 0.3\% BSA, 100 U of penicillin/ml, 100  \si{\micro}g of streptomycin/ml, and 100 \si{\micro}g of calcium chloride/ml).
At 48 hours post-transfection we harvested the virus-containing supernatant, pelletted cellular material by centrifugation at 300 x g's for 4 minutes, and stored aliquots at of the clarified viral supernatant -80 \si{\degreeCelsius }. 
We then titered thawed aliquots of viral by TCID50 on MDCK-SIAT1 cells, computing titers via the formula of \citet{reed1938simple}.
To generate our ``high-purity'' stocks of viruses for the single-cell sequencing experiments, we then infected MDCK-SIAT1 cells at an MOI of 0.01, and let the virus replicate for 36 hours prior to harvesting aliquots that were again clarified by low-speed centrifugation, aliquoted, stored at -80  \si{\degreeCelsius }, and titered by TCID50.
The high-MOI passage (high-defective particle) stock used in Figure~\ref{fig:viruspopulations} was generated by instead passaging in MDCK-SIAT1 cells twice at an MOI of 1 for 48 hours.

For the experiments in Figure~\ref{fig:coexpression}-Figure~supplement~\ref{figsupp:flowcyto}, we created a virus that carried an HA gene segment in which GFP replaced most of the HA coding sequence, following a scheme first described by \citet{marsh2007specific}.
Briefly, we created a plasmid encoding a viral RNA with GFP in place of the HA coding sequence in the context of the pHH21~\citep{Neumann:1999ws} reverse-genetics plasmid, removing potential start codons upstream of the GFP (see Figure~\ref{fig:coexpression}-source~data~\ref{figdata:HAflank} for the sequence of the viral RNA).
We then generated GFP-carrying virus by reverse-genetics in cells constitutively expressing HA~\citep{Doud:2016gm}.
To obtain sufficient titers, this HA-eGFP virus was expanded for 44 rather than 36 hours after initiating infection at an MOI of 0.01.

\subsection{qPCR}
For the qPCR in Figure~\ref{fig:viruspopulations} and Figure~\ref{fig:fluburdenbyflugene}-Figure~supplement~\ref{figsupp:cyclohexamide}, A549 cells were seeded at 3$\times$10\textsuperscript{5} cells / well in a 6-well tissue culture plate in D10 the day prior to infection. 
On the day of infection, a single well was trypsinized and the cells were counted in order to determine the appropriate amount of virus to use to achieve the intended MOI.
Immediately before infection, D10 was replaced with influenza growth media.
For cells incubated with cyclohexamide, the compound was added to a final concentration of 50  \si{\micro}g/ml at the time of infection -- previously confirmed to be sufficient to halt viral protein production~\citep{Killip:2014kz}.
RNA was purified using the QIAGEN RNeasy plus mini kit following manufacturer's instructions. 
cDNA was synthesized using an oligoDT primer and the SuperScript\textsuperscript{TM} III first-strand synthesis supermix from ThermoFisher using manufacturer's directions. 
Transcript abundance was measured using SYBR\textsuperscript{TM} green PCR master mix, using a combined anneal/extension step of 60 \si{\degreeCelsius } for one minute with the following primers: \emph{HA}: 5'-GGCCCAACCACACATTCAAC-3', 5'-GCTCATCACTGCTAGACGGG-3', \emph{IFNB1}: 5'-AAACTCATGAGCAGTCTGCA-3', 5'-AGGAGATCTTCAGTTTCGGAGG-3', \emph{L32}: 5'-AGCTCCCAAAAATAGACGCAC-3', 5'-TTCATAGCAGTAGGCACAAAGG-3'.   
Biological triplicates were performed for all samples.

For the measurements of viral genomic HA content in Figure~\ref{fig:viruspopulations}A, vRNA was harvested from 80 \si{\micro}l of viral supernatant by the addition of 600 \si{\micro}l of RLT plus before proceeding with the standard QIAGEN RNeasy Plus Mini kit protocol.
The cDNA was generated using SuperScript\textsuperscript{TM} III first-strand synthesis supermix using the manufacturer's protocol, and using the universal vRNA primers of \citet{Hoffmann:2001vj} with the modifications described in  \citet{Xue:2017dl}.
The qPCR was then performed as for mRNA measurements.
A standard curve was generated from three independent dilutions of the HA-encoding reverse genetics plasmid. 
All vRNA values represent three independent RNA extractions with two replicate qPCR measurements.

\subsection{Flow cytometry titering and analyses}
To determine viral titers in terms of HA-expressing units and for the flow cytometry in Figure~\ref{fig:coexpression}-Figure~supplement~\ref{figsupp:flowcyto}, A549 cells were seeded in a 6-well plate and infected as described above for the qPCR analyses.
Cells were harvested by trypsinization, resuspended in phosphate-buffered saline supplemented with 2\% heat-inactivated FBS, and stained with 10 \si{\micro}g/ml of H17-L19, a mouse monoclonal antibody confirmed to bind to WSN HA in a prior study~\citep{Doud:2017bw}.
After washing in PBS plus 2\% FBS, the cells were stained with a goat anti-mouse IgG antibody conjugated to APC.
Cells were then washed, fixed in 1\% formaldehyde, and washed further before a final resuspension and analysis. 
We then determined the fraction of cells that were HA positive and calculated the HA-expressing units assuming a Poisson infection process.

For Figure~\ref{fig:coexpression}-Figure~supplement~\ref{figsupp:flowcyto}, after gating to exclude multiplets in FlowJo, data were extracted using the R package flowCore~\citep{LeMeur:2007uo} and analyzed using a custom Python script.
Guassian kernel density estimates were obtained using the scipy stats package method, guassian\_kde, using automatic bandwidth determination~\citep{vanderWalt:2017dp}.

\subsection{Infections for single-cell mRNA sequencing}
Single-cell sequencing libraries were generated using the 10x Chromium Single Cell 3' platform~\citep{zheng2017massively} using the V1 reagents.

All time points except for the 8-hour replicate were prepared on the same day.
For the infections, A549 cells were seeded in a 6-well plate, with two wells per time point. 
A single well of cells was trypsinized and counted prior to initiation of the experiment for the purposes of calculating MOI.
Wild-type and synonymously barcoded virus were mixed to an estimated ratio of 1:1 based on prior, exploratory, single-cell analyses (data not shown). 
At the initiation of our experiment, the wells for all time points were changed from D10 to influenza growth media.
Cells were then infected with 0.3 HA-expressing units of virus per cell (a determined by flow cytometry).
The infections were performed in order of time point: first the 10-hour time point, then the 8-hour, and then the 6-hour time point.
At one hour after infection, the media for each time point was changed to fresh influenza growth media.
Note that the HA-expressing units were calculated without this additional washing step, and so likely represent an overestimate of our final infectious dose (consistent with the fact that fewer than 30\% of cells appear infected in the single-cell sequencing data).
All cells were then harvested for single-cell analysis concurrently -- ensuring all had spent equivalent time in changed media .
For our replicate 8-hour time point, cells were infected as above except that the cells were infected at 0.1 HA-expressing units of virus per cell but no wash step was performed, and the sample was prepared on a different day.
After harvest, cells were counted using disposable hemocytometers and diluted to equivalent concentrations with an intended capture of 3000 cells/sample following the manufacturer's provided by 10x Genomics for the Chromium Single Cell platform.
All subsequent steps through library preparation followed the manufacturer's protocol.
Samples were sequenced on an Illumina HiSeq. 
The deep sequencing data are available on the Sequence Read Archive under accession \jdbcomment{needs to be added.}

\subsection{Computational analysis of single-cell mRNA sequencing data}
Jupyter notebooks that perform all of the computational analyses are available at \url{https://github.com/jbloomlab/flu_single_cell} \jdbcomment{This GitHub repository will be made public upon acceptance of the manuscript. If you are a reviewer who needs access, please contact the editor.}, and are also available in Supplementary~file~\ref{suppfile:code}.

Briefly, the raw deep sequencing data were processed using the 10X Genomics software package CellRanger (version 2.0.0).
The reads were aligned to a concatenation of the human and influenza virus transcriptomes.
The human transcriptome was generated by filtering genome assembly GRCh38 for protein coding genes defined in the GTF file GRCh38.87.
The influenza virus transcriptome was generated from the reverse-complement of the wildtype WSN viral RNA sequences as encoded in the reverse-genetics plasmids (Figure~\ref{fig:workflow}-source~data~\ref{figdata:viralsequences}).
CellRanger calls cells based on the number of observed cell barcodes, and creates a cell-gene matrix.
We used custom Python code to annotate the cells in this matrix by the number of viral reads that could be assigned to the wildtype and synonymously barcoded virus.
This annotated cell-gene matrix is in Supplementary~file~\ref{suppfile:cellgenematrix}.
A Jupyter notebook that performs these analyses is in Supplementary~file~\ref{suppfile:code}.

The annotated cell-gene matrix was analyzed in R, primarily using the Monocle package (version 2.4.0)~\citep{Qiu:2017dw,Trapnell:2014kg}.
A Jupyter notebook that performs these analyses is in Supplementary~file~\ref{suppfile:code}.
For each sample, cell barcodes that had $>$2.5-fold fewer or more UMI counts mapping to cellular transcripts than the sample mean were excluded from downstream analyses (see red vertical lines in Figure~\ref{fig:cells}C).
In order to determine an appropriate cutoff for how many reads a cell needed to contain in order to be classified as infected, we calculated the mean viral barcode purity across all cells that contained at least a given fraction of viral mRNA and had multiple viral reads that could be assigned a barcode (Figure~\ref{fig:viralbarcodes}B,C).
We then determined the threshold fraction of viral mRNA at which the mean purity no longer increased as a function of the amount of viral mRNA.
This threshold represents the point at which we have effectively eliminated cells that have low barcode purity simply due to lysis-acquired reads sampled randomly from both viral barcodes.
The 10-hour sample has the most viral mRNA, and so is expected to have the most lysis-associated influenza reads.
We therefore determined the threshold using just this sample (Figure~\ref{fig:viralbarcodes}C), and then applied this threshold to all samples.
This procedure is expected to be conservative, and may miss some truly infected cells with very low amounts of viral mRNA.
For subsequent analyses, we retained all infected cells and a subsample of uninfected cells (the greater of 50 or the number of infected cells for that sample).
The rationale for subsampling the uninfected cell is that the vast majority of cells are uninfected, and we did not want these cells to completely dominate the downstream analyses.

For the semi-supervised t-SNE clustering, we used Monocle's cell hierarchy function to bin cells into those with no viral mRNA, <2\% viral mRNA, between 2\% and 20\% viral mRNA, and >20\%. 
Candidate marker genes for t-SNE dimensionality reduction were then determined using the Monocle function markerDiffTable, excluding the effects of sample variation and the number of genes identified in a given cell, using a q-value cutoff of 0.01.
The specificity of these markers was determined using the function calculateMarkerSpecificity -- the top 40 markers were retained, and used to place populations in a two-dimensional plane based on tSNE dimensionality reduction.

For the analyses of cellular genes that differed in expression as a function of the amount of viral mRNA, we only considered cells that expressed all 8 viral mRNAs to avoid effects driven simply by viral gene absence.
We also only considered cellular genes in the differential gene analysis, since viral gene expression will tautologically co-vary with the amount of viral mRNA.
Additionally, because influenza virus has the capacity to degrade or prevent the synthesis of host mRNAs~\citep{BercovichKinori:2016iw} and contributes significantly to the total number UMIs in some cells, we calculate size factors (a scalar value representing efficiency of UMI capture) based on cellular transcripts alone. 
Finally, we assigned all cells a ceiling fraction of mRNA from virus of 25\% so that a few extremely high-expressing cells did not dominate. 
Cellular genes with expression that co-varied with the fraction of viral mRNAs in a cell were then determined using the Monocle differentialGeneTest, after removing variance explained by sample to sample variation. 
Figure~\ref{fig:cellulargenes} shows all genes that were significantly associated with the fraction of mRNA from virus at a false discovery rate of 0.1.

\section{Acknowledgments}
We thank the Fred Hutch Genomics Core for performing the Illumina deep sequencing.
